---
title: "Variations in American Crow Densities"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
```

###Selecting the PSUs for Rossmoor
```{r}
set.seed(152)
rossmoor_samp <- sample(1:220, 15, replace = FALSE)
rossmoor_samp
```

###Selecting PSUs for Lucerne Valley
```{r}
set.seed(114)
lucerne_samp <- sample(1:216, 15, replace = FALSE)
lucerne_samp
```

#Load the Data
```{r}
crow <- read.csv("crowdata.csv")
```

```{r, include= FALSE}
head(crow)
names(crow) <- c("Grid", "Length", "Width", "Crows", "Time")
```

```{r}
head(crow)
```

```{r}
crow$Length <- crow$Length * 1609.34
crow$Width <- crow$Width * 1609.34
```

#Analysis for Lucerne Valley
##Total Estimates
```{r}
lucerne <- crow[str_detect(crow$Grid, "LV"), ]
head(lucerne)
```

```{r}
#Estimate of the Total Crows in Lucerne Valley
NL <- 220
nL <- 15
lucerne_total <- (NL / nL) * sum(lucerne$Crows)
lucerne_total
```

```{r}
#Variance of Total Crow Estimate in Lucerne Valley
lucerne_sample_var <- (1 / (nL - 1)) * sum((lucerne$Crows - (lucerne_total / NL)) ** 2)
lucerne_var <- (NL ** 2) * (1 - (nL / NL)) * lucerne_sample_var / nL

lucerne_var
```

```{r}
#Standard Error of Total Crow Estimate in Lucerne Valley
lucerne_se <- sqrt(lucerne_var)
lucerne_se
```

```{r}
#Standard Deviation of Total Crow Estimate in Lucerne Valley
lucerne_sd <- lucerne_se * sqrt(nrow(lucerne))
lucerne_sd
```

```{r}
#95% Confidence Interval for Total Crows in Lucerne Valley
c(lucerne_total - (1.96 * lucerne_se), lucerne_total + (1.96 * lucerne_se))
```

##Density Estimates
```{r}
#Estimate of Crow Density in Lucerne Valley
lucerne_density <- lucerne_total / 4.25
lucerne_density
```

```{r}
#Variance of Crow Density in Lucerne Valley
lucerne_density_sample_var <- (1 / (nL - 1)) * sum((lucerne$Crows - (lucerne_total / NL)) ** 2) * (1 / (4.25 ** 2))
lucerne_density_var <- (NL ** 2) * (1 - (nL / NL)) * lucerne_density_sample_var / nL

lucerne_density_var
```

```{r}
#Standard Error of Crow Density in Lucerne Valley
lucerne_density_se <- sqrt(lucerne_density_var)
lucerne_density_se
```

```{r}
#Standard Deviation of Crow Density in Lucerne Valley
lucerne_density_sd <- lucerne_density_se * sqrt(nrow(lucerne))
lucerne_density_sd
```

```{r}
#95% Confidence Interval for Total Crow Density in Lucerne Valley
c(lucerne_density - (1.96 * lucerne_density_se), lucerne_density + (1.96 * lucerne_density_se))
```

#Anaylsis for Rossmoor, California
##Total Estimates
```{r}
rossmoor <- crow[str_detect(crow$Grid, "R"), ]
head(rossmoor)
```

```{r}
#Estimate of the Total Crows in Rossmoor
NR <- 216
nR <- 15
rossmoor_total <- (NL / nL) * sum(rossmoor$Crows)
rossmoor_total
```

```{r}
#Variance of Total Crow Estimate in Rossmoor
rossmoor_sample_var <- (1 / (nR - 1)) * sum((rossmoor$Crows - (rossmoor_total / NR)) ** 2)
rossmoor_var <- (NR ** 2) * (1 - (nR / NR)) * rossmoor_sample_var / nR

rossmoor_var
```

```{r}
#Standard Error of Total Crow Estimate in Rossmoor
rossmoor_se <- sqrt(rossmoor_var)
rossmoor_se
```

```{r}
#Standard Deviation of Total Crow Estimate in Rossmoor
rossmoor_sd <- rossmoor_se * sqrt(nrow(rossmoor))
rossmoor_sd
```

```{r}
#95% Confidence Interval for Total Crows in Rossmoor
c(rossmoor_total - (1.96 * rossmoor_se), rossmoor_total + (1.96 * rossmoor_se))
```

##Denisty Estimates
```{r}
#Estimate of Crow Density in Rossmoor
rossmoor_density <- rossmoor_total / 4.25
rossmoor_density
```

```{r}
#Variance of Crow Density in Rossmoor
rossmoor_density_sample_var <- (1 / (nR - 1)) * sum((rossmoor$Crows - (rossmoor_total / NR)) ** 2) * (1 / (4.25 ** 2))
rossmoor_density_var <- (NR ** 2) * (1 - (nR / NR)) * rossmoor_density_sample_var / nR

rossmoor_density_var
```

```{r}
#Standard Error of Crow Density in Rossmoor
rossmoor_density_se <- sqrt(rossmoor_density_var)
rossmoor_density_se
```

```{r}
#Standard Deviation of Crow Density in Rossmoor
rossmoor_density_sd <- rossmoor_density_se * sqrt(nrow(rossmoor))
rossmoor_density_sd
```

```{r}
#95% Confidence Interval for Total Crow Density in Rossmoor
c(rossmoor_density - (1.96 * rossmoor_density_se), rossmoor_density + (1.96 * rossmoor_density_se))
```

#Comparison of the Two Groups

###Null Hypothesis: The density of crows in Lucerne Valley, California, is less than or equal to the density of crows in Rossmoor, California.
###Alternative Hypothesis: The density of crows in ZLucerne Valleu, California, is larger than the density of crows in Rossmoor, California.

```{r}
#Test Statistic for the Difference in Densities
difference <- lucerne_density - rossmoor_density

variance_of_difference <- (rossmoor_density_sample_var / nR) + (lucerne_density_sample_var / nL)

test_stat <- difference / sqrt(variance_of_difference)
```

```{r}
#P-value for the Calculated Test Statistic
p_value <- pt(-abs(test_stat), df = 14)
p_value
```













